Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-03-13 manuscript/manuscript.Rmd
assorted edits, plot fix

---
title: "Estimating the likelihood of student replication success after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: TODO x a lot
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt
header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

::: {.abstract data-latex="" lang="en"}
<!-- In order to do cumulative science, results need to be replicable. However, an individual direct replication could get a different result than an original ("replication failure") for a number of reasons, which vary in whether they would apply to a potential second replication. Given one failed replication, how likely is a second attempt at replication to succeed?-->

As a pragmatic matter, when a replication fails, scientists have to decide: do they make a second attempt at replication, or move on?
Given the relatively low rates of replication success reported in large scale replication projects, this choice point is one facing many scientists in psychology who attempt to replicate a study before building on it.
Here, we report on 17 re-replications of experiments with failed replicated reported in Boyce et al. (2023).
In 5/17 of these rescue projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes.
While a few studies with failed replications were rescued with larger sample sizes and experimental tweaks, our overall results indicate that most failed replications stayed failed.
If the goal is to do cumulative science, one may wish to stop after one failed replication.
:::

```{=html}
<!-- Other Title ideas:

Can failed replications be rescued? Evidence from N re-replication experiments suggests rarely.

Try again? Attempts to rescue failed replications were generally unsuccessful

A low rate of successfully replicating experiments with a prior unsuccessful replication

Once is probably enough. Low replication rates when re-replicating N studies with failed replications.

* may want to include "student" or "class"

Estimating the replicability of experiments with a prior unsuccessful replication

Estimating the likelihood of [student] replication success after an initial failure to replicate
-->
```
# Introduction

@@

@@

TODO DISCUSSION OF THE LITERATURE What these replication rates say about the truth of the literature is beyond our scope, but these rates do indicate that scientists are far from guaranteed at being able to replicate effects from the literature in their own labs.
Thus, the question of what to do with a failed replication is a practical one that comes up frequently.

@@

While one can theorize after the fact about potential causes of replication failure, in most circumstances it is hard to diagnose for certain what made for the difference in the results.
An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error).
Re-replication could serve as a way of triangulating on reasons for replication failure, as well as another chance for a scientist to get a working protocol that picks up the effect of interest.
However, re-replication has not been studied much, so it is uncertain how definitive a replication failure is.
That is, conditioned on a failed replication, how likely is a re-replication to succeed?
Re-replication was tested in a sample of 10 psychology studies in @ebersole2020.
After @openscienceconsortium published their results with a replication rate of TODO, @gilbert2016 raised concerns that methodological differences were the problem.
In 11 of the OSF studies, concerns about protocol fidelity had been raised by original authors prior to data-collection (and thus not contingent on results).
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication [@openscienceconsortium].
In a follow-up, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts, thus testing whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result: 0 of the RP:P protocols found significant results, 2 of the 10 revised protocols did (but not the 1 that had a significant result in the original RP:P sample).
The replication effect size was much smaller CITE NUMBER [@ebersole2020].
We could frame @ebersole2020 as an attempt to rescue the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks.
This suggests a successful re-replication rate of 2/9 (22%) for studies with non-significant first replications, under fairly favorable, high resource circumstances.

@@

Factors that may be fixable include sample size issues with the replication, assuming the re-replication can recruit a larger sample or reduce attrition.
It may be possible to fix underpoweredness and inflated effect size, at least to an extent.
If the effect size was inflated and the first replication powered for the effect, a better powered (but not huge) replication may be able to recover a smaller, but still convincing effect.
Failures due to differences between the materials may or may not be fixable.
If the original materials are still unavailable, a second replication is no better off.
If a second replication can get closer to the original, either by using closer materials, or better adapting the study to a new environment, the rescue might succeed at better approximating the results of the original.
A failed manipulation check failure may not be fixable, but it might at least be diagnoseable.

@@

In the present paper, we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023.
This is a non-random sample, and given the small number of projects, we provide descriptive statistics and discuss qualitatively some potential sources of differential re-replication outcomes.
[possibly drop last sentence?]

@@

A spreadsheet of projects, individual project write-ups (both replications and rescues), links to individual project data and analyses for rescue projects, and the analytic code for this paper are all available at TODO OSF LINK GOES HERE.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (to ensure we could afford to increase the sample size).
We then contacted the replication project authors for permission to share their report and github repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects.
`r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of the rescue paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the original replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies.
For instance, the rescue of @paxton2012 switched from using the CRT which is overused in online samples for the newer TODO.
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli.
@tarampi2016 original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with press keyboard keys to indicate.
TODO others to note?
Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

@@

## Pre-registration

We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects.
Prediction intervals depend on both the original effect size and variance CITATION TODO, and also the variance of the comparison (replication) study.
Thus we cannot show *the* prediction interval for the original study, but would have to show a prediction interval between each pair of studies, which we thought would not offer clarity.

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP.
Disagreements were resolved through discussion.
As a compliment to the subjective rating of overall success, we stastically compared one key measure of interest for each study, following @boyce2023. In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, replication, and re-replication.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, replication, and rescue (these were already rated for original and replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

@@

Across the `r nrow(subj)` replications, `r nrow(success)` mostly or fully replicated the original results according to the subjective replication ratings.
`r pluck(tallied,2,1)` had a rating of 0, `r pluck(tallied, 2,2)` got a rating of .75, and `r pluck(tallied, 2,3)` got a rating of 1.
Thus, a first pass answer to the question "how often can a failed replication be salvaged?" is `r round(nrow(success)/nrow(subj)*100)`% (bootstrapped 95% CI: `r round(bootstrap[2]*100)`% - `r round(bootstrap[3]*100)`%) of the time.

In the original replication sample from @boyce2023, `r boyce2023_sample |> filter(sub_rep>.5) |> nrow()` out of `r boyce2023_sample |> nrow()` replications (`r round(percent_rep*100)`%) mostly or fully replicated (i.e. received a subjective replication score of .75 or 1).
Note that @boyce2023 report the average replication score as a percent success (49%), but given that we considered studies with a subjective score of .5 as eligible to be rescued, we recomputed the success rate when scores of 0-.5 as failures and .75 and 1 as successes.
If the re-replication rate in our sample is representative of the re-replication rate for the initially non-replication studies, then the combined chance of mostly or fully replicating in a first replication or a follow-up replication is `r round(overall*100)`%.

Another successful replication was that of @ngo2019.
Here, the original study found a large effect.
The first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people.
It failed to find the effect.
The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separate reasons that slight changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.

library(kableExtra)
knitr::kable(cors, digits = 3, align='rcc', format="latex", booktabs=T, linesep="", caption="Correlations between an individual predictor and the subjective replication score of the rescue project.")

We ran correlations between the set of predictor variables used in @boyce2023 and the subjective replication scores of the rescues.
We also added some (non-preregistered) predictors related to the sample size of the first replication, after seeing the successful re-replications of @ngo2019 and @krauss2003.

All of the correlations are presented in Figure \@ref(tab:cor-table).
As the number of rescues is small, and many of these predictors are correlated, we caution against over-interpretation.
The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of these effects meet the conventional significance threshold.

Small replication samples relative to original and rescue could be due to both a) powering a replication according to a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample.
Since relative sizes of the studies may play a role in replication success and how probative replications are, we show the sample sizes in Table \@ref(tab:tab-size).

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original.
In the rescue projects, we aimed to have methods be as close as was feasible or appropriate.
(As discussed above, changes in time, population, or setting may necessitate adaptation of a study to maximize fidelity.) However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools.
Table \@ref(tab:tab-size) shows the closeness of each re-replication according to the classification scheme from @lebel2018.

```{r smd, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

#test on just one
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")
shape <- c("original"=21,"rep1"=22,"rescue"=23,"additional"=24)

big <- parsed_d |> filter(type=="original") |> filter(calc_d_calc>1) |> select(target_lastauthor_year)
for_plotting_d <- parsed_d |> filter(!is.na(calc_d_calc)) |> 
  mutate(point=calc_d_calc, low=calc_d_calc-1.96*calc_d_calc_se, high=calc_d_calc+1.96*calc_d_calc_se)

@@

smd_scale_big <- ggplot(for_plotting_d |> inner_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

@@

plot_grid(smd_scale_big, smd_scale_small, rel_heights=c(.3,1), nrow=2)
```

```{r , include=F, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

#test on just one
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")
shape <- c("original"=21,"rep1"=22,"rescue"=23,"additional"=24)

smd_scale <- ggplot(for_plotting_d,aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

smd_scale

As a complement to the holistic, qualitative replication ratings used above, we also statistically compared the effect sizes of the rescue, original replication, and original study on one key measure per study.
We followed @boyce2023 in determining a key measure for each study.
When we were aware of additional direct replications (either from other class projects, or external replications in the literature), we also consider the effect sizes obtained in these additional replications.

@@

[Ben: I think the paragraph below could use more explanation about what $p$-original is, but I'm not confident enough in my understanding of it to write it myself.]

@@

```{r, include=F,  fig.caption="Plot of the prediction intervals of the original studies compared to replication results. The open circles are SMD effect sizes from the original studies, error bars are the prediction intervals, and filled shapes denote replication effect sizes. Prediction intervals in green contain the replication effect, those in grey do not. The largest effect sizes are separated in a separate panel to allow better visualization. "}
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")
shape <- c("original"=21,"rep1"=22,"rescue"=23,"additional"=24)
col_code <- c("TRUE"="darkgreen", "FALSE"="black")

big <- predInts_het |> filter(d>1) |> select(target_lastauthor_year)

small_predInt <-  ggplot(predInts_het |> filter(!is.na(d_inside)) |> anti_join(big), aes(x=reorder(name_pretty, d, FUN=max), y=rep_d, ymin=d_low, ymax=d_high, fill=type, color=d_inside, group=desc(type), shape=type))+
  geom_errorbar(size=.5, width=.25, position=position_dodge(width=.4), alpha=.5)+
  geom_point(aes(y=d), position=position_dodge(width=.4), shape=1, alpha=.5)+
     geom_point(size=2.5, position=position_dodge(width=.4), color="transparent")+
  coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position="bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_fill_manual(values=colors)+
  scale_shape_manual(values=shape)+
  scale_color_manual(values=col_code)+
  labs(y="predInts on SMD scale", x="")

big_predInt <-  ggplot(predInts_het |> filter(!is.na(d_inside)) |> inner_join(big), aes(x=reorder(name_pretty, d, FUN=max), y=rep_d, ymin=d_low, ymax=d_high, color=d_inside, fill=type, group=desc(type), shape=type))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4), alpha=.5)+
  geom_point(aes(y=d), position=position_dodge(width=.4), shape=1, alpha=.5)+
     geom_point(size=2.5, position=position_dodge(width=.4), color="transparent")+
  coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position="none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_fill_manual(values=colors)+
  scale_shape_manual(values=shape)+
  scale_color_manual(values=col_code)+
  labs(y="predInts on SMD scale", x="")

plot_grid(big_predInt, small_predInt, nrow=2, rel_heights = c(.3,1))

@@

>>>>

---
title: "Estimating the replicability of psychology experiments after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Psych
    footnote:
      - corresp
  - name: Ben Prystawski
    affiliation: Psych
  - name: Adani Abutto
    affiliation: Psych
  - name: Emily Chen
    affiliation: Psych
  - name: Ziwen Chen
    affiliation: GSB
  - name: Howard Chiu
    affiliation: Edu
  - name: Irmak Ergin
    affiliation: Psych
  - name: Anmol Gupta
    affiliation: Psych
  - name: Chuqi Hu
    affiliation: Symsys
  - name: Bendix Kemmann
    affiliation: Philosophy
  - name: Nastasia Klevak
    affiliation: Psych
  - name: Verity Y. Q. Lua
    affiliation: Psych
  - name: Mateus M. Mazzaferro
    affiliation: Edu
  - name: Khaing Mon
    affiliation: Symsys
  - name: Dan Ogunbamowo
    affiliation: Psych
  - name: Alexander Pereira
    affiliation: Philosophy
  - name: Jordan Troutman
    affiliation: CS
  - name: Sarah Tung
    affiliation: Psych
  - name: Raphael Uricher
    affiliation: Psych
  - name: Michael C. Frank
    affiliation: Psych
address:
  - code: Psych
    address: Department of Psychology, Stanford University
  - code: GSB
    address: Graduate School of Business, Stanford University
  - code: Edu
    address: Graduate School of Education, Stanford University
  - code: Symsys
    address: Symbolic Systems Program, Stanford University
  - code: Philosophy
    address: Department of Philosophy, Stanford University
  - code: CS
    address: Department of Computer Science, Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt
header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

::: {.abstract data-latex="" lang="en"}
When a replication fails, scientists have to decide whether to make a second attempt or move on. Researchers who attempt to replicate studies before building on them often face this decision, given the relatively low rates of replication success in psychology. Here, we report on 17 re-replications of experiments with failed replications reported in Boyce et al. (2023). In 5/17 of these "rescue" projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few replications were rescued with larger sample sizes and minor methodological changes, most re-replications failed again. In many situations, it may be most efficient to stop pursuing an effect after a single failed replication.
:::
TODO FIX CITATION / BIB INFO

# Introduction

@@

____
TODO VB IS UNSURE WHAT EXACTLY THE PROPOSED REORDERING IS

@@

 What these replication rates say about the truth of the literature is beyond our scope, but these rates do indicate that scientists are far from guaranteed at being able to replicate effects from the literature in their own labs.
Thus, the question of what to do with a failed replication is a practical one that comes up frequently.

@@

While one can theorize after the fact about potential causes of replication failure, in most circumstances it is hard to diagnose for certain what made for the difference in the results.
An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error).
Re-replication could serve as a way of triangulating on reasons for replication failure, as well as another chance for a scientist to get a working protocol that picks up the effect of interest.
However, re-replication has not been studied much, so it is uncertain how definitive a replication failure is.
That is, conditioned on a failed replication, how likely is a re-replication to succeed?
Re-replication was tested in a sample of 10 psychology studies in @ebersole2020.
After @openscienceconsortium published their results with a replication rate of TODO, @gilbert2016 raised concerns that methodological differences were the problem.
In 11 of the OSF studies, concerns about protocol fidelity had been raised by original authors prior to data-collection (and thus not contingent on results).
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication [@openscienceconsortium].
In a follow-up, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts, thus testing whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result: 0 of the RP:P protocols found significant results, 2 of the 10 revised protocols did (but not the 1 that had a significant result in the original RP:P sample).
The replication effect size was much smaller CITE NUMBER [@ebersole2020].
We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks.
This suggests a successful re-replication rate of 2/9 (22%) for studies with non-significant first replications, under fairly favorable, high resource circumstances.

@@

Factors that may be fixable include sample size issues with the replication, assuming the re-replication can recruit a larger sample or reduce attrition.
It may be possible to fix underpoweredness and inflated effect size, at least to an extent.
If the effect size was inflated and the first replication powered for the effect, a better powered (but not huge) replication may be able to recover a smaller, but still convincing effect.
Failures due to differences between the materials may or may not be fixable.
If the original materials are still unavailable, a second replication is no better off.
If a second replication can get closer to the original, either by using closer materials, or better adapting the study to a new environment, the second replication might succeed at better approximating the results of the original.
A failed manipulation check failure may not be fixable, but it might at least be diagnoseable.

@@

In the present paper, we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023.
TODO PREVIEW RESULTS??

@@

A spreadsheet of projects, individual project write-ups (both first replications and rescues), links to individual project data and analyses for rescue projects, and the analytic code for this paper are all available at TODO OSF LINK GOES HERE.

We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance CITATION TODO, and also the variance of the comparison (replication or rescue) study.
Thus we cannot show *the* prediction interval for the original study, but would have to show a prediction interval between each pair of studies, which we thought would not offer clarity.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (to ensure we could afford to increase the sample size).
We then contacted the replication project authors for permission to share their report and github repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects.
`r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of this paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies.
For instance, the rescue of @paxton2012 switched from using the CRT which is overused in online samples for the newer TODO.
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli.
@tarampi2016 original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with pressing keyboard keys to indicate the direction of the a turn.
TODO others to note?
Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

@@

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP.
Disagreements were resolved through discussion.
As a compliment to the subjective rating of overall success, we stastically compared one key measure of interest for each study, following @boyce2023. In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, first replication, and rescue.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, first replication, and rescue (these were already rated for original and first replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

@@

Across the `r nrow(subj)` rescue projects, `r nrow(success)` mostly or fully replicated the original results according to the subjective replication ratings.
`r pluck(tallied,2,1)` had a rating of 0, `r pluck(tallied, 2,2)` got a rating of .75, and `r pluck(tallied, 2,3)` got a rating of 1.
Thus, a first pass answer to the question "how often can a failed replication be salvaged?" is `r round(nrow(success)/nrow(subj)*100)`% (bootstrapped 95% CI: `r round(bootstrap[2]*100)`% - `r round(bootstrap[3]*100)`%) of the time.

In the original replication sample from @boyce2023, `r boyce2023_sample |> filter(sub_rep>.5) |> nrow()` out of `r boyce2023_sample |> nrow()` replications (`r round(percent_rep*100)`%) mostly or fully replicated (i.e. received a subjective replication score of .75 or 1).
Note that @boyce2023 report the average replication score as a percent success (49%), but given that we considered studies with a subjective score of .5 as eligible to be rescued, we recomputed the success rate when scores of 0-.5 as failures and .75 and 1 as successes.
If the re-replication rate in our sample is representative of the re-replication rate for the initially non-replicating studies, then the combined chance of mostly or fully replicating in a first replication or one follow-up replication is `r round(overall*100)`%.

```{r smd, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. In a few cases, the first replication's key effect was non-zero in the same direction as the original; however, in these cases the larger pattern of results was not fully consistent between original and first replication.  "}

#test on just one
colors <- c("Original study"="black", "First replication"="#377EB8", "Rescue"="#E41A1C", "Additional replication"="#984EA3")
shape <- c("Original study"=21,"First replication"=22,"Rescue"=23,"Additional replication"=24)

big <- parsed_d |> filter(type=="original") |> filter(calc_d_calc>1|target_lastauthor_year=="yeshurun2003") |> select(target_lastauthor_year)
for_plotting_d <- parsed_d |> filter(!is.na(calc_d_calc)) |> 
  mutate(point=calc_d_calc, low=calc_d_calc-1.96*calc_d_calc_se, high=calc_d_calc+1.96*calc_d_calc_se) |> 
  mutate(type=case_when(
    type=="original" ~ "Original study",
    type=="rep1" ~ "First replication",
    type=="rescue" ~ "Rescue",
    type=="additional" ~ "Additional replication"
  ),
  type=factor(type, levels=c("Original study", "First replication", "Rescue", "Additional replication")))

@@

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

@@

plot_grid(smd_scale_big, smd_scale_small, rel_heights=c(.4,1), nrow=2)
```

As a complement to the holistic, qualitative replication ratings used above, we also statistically compared the effect sizes of the rescue, first replication, and original study on one key measure per study.
We followed @boyce2023 in determining a key measure for each study.
When we were aware of additional direct replications (either from other class projects, or external replications in the literature), we also consider the effect sizes obtained in these additional replications.

@@

[TODO Ben: I think the paragraph below could use more explanation about what $p$-original is, but I'm not confident enough in my understanding of it to write it myself.]

@@

library(kableExtra)
knitr::kable(cors, digits = 3, align='rcc', format="latex", booktabs=T, linesep="", caption="Correlations between an individual predictor and the subjective replication score of the rescue project. The first set of predictors were pre-registered based on the correlates used in Boyce et al. (2023). The last three predictors were added post-hoc.") |> row_spec(11, hline_after=T)

@@

We ran correlations between the set of predictor variables used in @boyce2023 and the subjective replication scores of the rescues.
We also added some (non-preregistered) predictors related to the sample size of the first replication, after seeing the successful re-replications of @ngo2019 and @krauss2003, both studies with small replication samples.

All of the correlations are presented in Figure \@ref(tab:cor-table).
As the number of rescues is small, and many of these predictors are correlated, we caution against over-interpretation.
The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of the pre-registered correlates meet the conventional significance threshold; the two correlates based on ratios that reflect a relatively smaller replication sample size are marginally significant.

Small replication samples relative to original and rescue could be due to both a) powering a replication according to a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample.
Since relative sizes of the studies may play a role in replication success and how probitive replications are, we show the sample sizes in Table \@ref(tab:tab-size).

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original.
In the rescue projects, we aimed to have methods be as close as was feasible or appropriate.
(As discussed above, changes in time, population, or setting may necessitate adaptation of a study to maximize fidelity.) However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools.
Table \@ref(tab:tab-size) shows the closeness of each first replication and rescue according to the classification scheme from @lebel2018. TODO CLOSENESS NUMBERS SHOULD BE DOUBLE CHECKED AND FIRST REPS REVISED AS NEEDED

# Case studies

Another successful rescue was that of @ngo2019.
Here, the original study found a large effect.
The first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people.
It failed to find the effect.
The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separate reasons that slight changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.
